---
title: "Coding Problem Set 2"
author: "Keerthana K M 231057007"
date: "2023-08-26"
output: html_document
---

```{r}
#initially, the bus is empty at the first stop
onboard = numeric(10)
offboard = numeric(10)

for(i in c(1:10)){ #ten stops
  # there is a 50% chance of 0 passengers, or 40% chance of one passenger, or 10% chance of 2 passengers getting on-board
  onboard[i] = 0.5*0 + 0.4*1 + 0.1*2
}

#since people get on the bus at first stop, let us assume they start getting down from the 2nd stop and further
for(i in c(2:10)){
  # there is a 20% chance of people getting down the bus independently (1 passenger)
  offboard[i] = 0.2*1
}

# at first stop, no one gets down the bus, hence probability of a passenger getting down the bus at first stop is 0
offboard[1] = 0

```

```{r}
#0th passenger probability
passenger0 = function(){
  return(0.5)
}
```
```{r}
#1st passenger probability
passenger1 = function(){
  return(0.4)
}
```
```{r}
#2nd passenger probability
passenger2 = function(){
  return(0.1)
}
```

```{r}
passengers = function(){
  result = numeric(10)
  for(i in c(1:10)){
    result[i] = onboard[i] #process of on-boarding and off-boarding the bus for each passenger
  }
  return(result)
}
```

```{r}
nsims = length(c(4:1000)) #simulations for 1000-2 passengers
simulatedData = replicate(nsims, passengers())
str(simulatedData)
```

```{r}
onboarded_passengers = numeric(1000)
onboarded_passengers[1] = passenger0()
onboarded_passengers[2] = passenger1()
onboarded_passengers[3] = passenger2()
onboarded_passengers[4:1000] = passengers()
#str(onboarded_passengers)
```

```{r}
val = numeric(1000)
checkEvent = function(data){
  for(i in c(1:1000)){
    if(i==1){
      val[i] = onboarded_passengers[i]
    }else if(i==2){
      val[i] = onboarded_passengers[i]
    }else if(i==3){
      val[i] = onboarded_passengers[i]
    }else{
      val[i] = onboarded_passengers[i]
    }
  }
  return(val)
}
```

```{r}
passen = mean(checkEvent())*100 #Percentage probability of no. of passengers in the bus
print(passen)
```
There are about 60% of passengers in the bus during the ten stops. This means 40% of people have got down from the bus in the process.
Hence, the percentage probability of bus being empty after tenth stop is 
(no. of passengers in the bus - no. of passengers who got down)/total number of stops

```{r}
off = 100-passen
result = (passen - off)/11
print(result) #percentage probability of remaining no. of people getting down the bus by next stop, i.e. the 11th stop
```
Since the percentage is as low as ~2%, we can conclude that the bus will not be empty after the 10th stop.

```{r}
# Sampling space for leaving the bus
s1 = c(1, 0) # 1 corresponds to leaving the bus
p1 = c(0.2, 0.8)
p = array(c(replicate(3, c(0.8, 0.2)), replicate(4, c(0.2, 0.8)), replicate(3, c(0.8, 0.2))), dim = c(2, 10))

# Sampling space for boarding the bus
s2 = c(0, 1, 2)
p2 = c(0.5, 0.4, 0.1)
pass = array(c(replicate(1, c(0.5, 0, 0)), replicate(1, c(0, 0.4, 0)), replicate(1, c(0, 0, 0.1)), replicate(7, c(0.5, 0.4, 0.1))), dim = c(3, 10))
print(pass)

# Simulate the number of passengers from stop-1 to stop-10
nstops = 10 # number of stops
npassengers = 0 # number of passengers in the bus initially

simulatedData = replicate(nstops, pass)

mean(simulatedData)
```


